Constraint result to the self-interacting dark matter particle mass from the observational data

### 2025-06-02 article/article.tex
add cross-section figure contrast

\section{Introduction}

@@

   In this paper, we use the newly revised data from $H(z)$, supernovae Ia 
   and baryon acoustic oscillation by using the Markov chain Monte
   Carlo (MCMC) method. The result gives that the lower limit of the interacting
   dark matter particle mass is about 10 eV.

\section{The basic equations in the IDM model}

   We assume that the total density of the cosmic fluid obeys
   the collisional Boltzmann equation (\cite{Kolb:1990vq})
   \begin{equation}
      \dot{\rho}+3H\rho+\kappa\rho^2-2\Psi=0,
   \end{equation}
   where $\rho$ is the total energy-density of the cosmic fluid,
   containing dark matter, baryons, and any type of exotic energy,
   $\Psi$ is the rate of creation of DM particle pairs, and the
   annihilation parameter $\kappa(\geq0)$ is given by:
   \begin{equation}
      \kappa=\frac{\langle\sigma u\rangle}{M_x},\label{eq:2}
   \end{equation}
   where $\sigma$ is the cross-section for annihilation, $u$ is
   the mean particle velocity, and $M_x$ is the mass of the DM
   particle.
   Compared to the usual fluid equation, the effective pressure term
   is \begin{equation}
      P=\frac{\kappa\rho^2-\Psi}{3H}.
   \end{equation}
   When $\kappa\rho^2-\Psi<0,$ what means that the IDM particle
   creation term is larger than the annihilation item, IDM may serve
   as a negative pressure source in the global dynamics of the Universe,
   like the role of Dark Energy in the general cosmological models.

   \cite{Basilakos_2009} identified two functional forms for which
   the previous Boltzmann equation can be solved analytically.
   Refering to Appendix B in \cite{Basilakos_2009}, only one of these two is of interest
   because it provides a ``$\propto a^{-3}$'' dependence of the scale factor,
   which is \begin{equation}
      \Psi(a)=aH(a)R(a)=C_1(n+3)a^nH(a)+\kappa C_1^2a^{2m}.
   \end{equation}
   And the total energy density is
   \begin{equation}
      \rho(a)=C_1a^n+\frac{a^{-3}F(a)}{C_2-\int_1^{a}x^{-3}f(x)F(x)\mathrm{d}x},\label{eq:5}
   \end{equation}
   where $f(a)=-\kappa/[aH(a)],$ and the kernal function $F(a)$ has
   the form \begin{equation}
      F(a)=\exp\left[-2\kappa C_1\int_1^{a}\frac{x^{n-1}}{H(x)}\mathrm{d}x\right].
   \end{equation}
   The first term of Eq.(\ref{eq:5}) is the density corresponding to the
   residual matter creation that results from a possible disequilibrium
   between the particle creation and annihilation processes, while the second
   term can be viewed as the energy density of the self-IDM particles that are
   dominated by the annihilation process.

@@

   If $n=0$, the global density evolution can be transformed as
   \begin{equation}
      \rho(a)=C_1+a^{-3}\frac{e^{-2\kappa C_1(t-t_0)}}{C_2-\kappa Z(t)},
   \end{equation}
   where $Z(t)=\int_{t_0}^ta^{-3}e^{-2\kappa C_1(t'-t_0)}\mathrm{d}t'$ (\cite{Basilakos_2009}).
   Using the usual unit-less $\Omega$-like parameterization, we obtain that
   \begin{equation}
      \left(\frac{H}{H_0}\right)^2=\Omega_{1,0}+\frac{\Omega_{1,0}\Omega_{2,0}a^{-3}e^{-2\kappa C_1(t-t_0)}}
      {\Omega_{1,0}+\kappa C_1\Omega_{2,0}Z(t)},\label{eq:8}
   \end{equation}
   where $\Omega_{1,0}=8\pi GC_1/3H_0^2$ and $\Omega_{2,0}=8\pi G/3H_0^2C_2$,
   which related to $\Omega_{\Lambda}$ and $\Omega_m$ in the $\Lambda$CDM model,
   respectively.
   From Eq.({\ref{eq:2}}), we can also give the mass of the DM particle
   related to the range of $\kappa C_1$(in the unit of Gyr${}^{-1}$)
   \begin{equation}
      M_x=\frac{3.325\times10^{-12}}{\kappa C_1}
      \frac{\langle\sigma u\rangle}{10^{-23}}h^2(1-\Omega_{2,0})\,\text{GeV},\label{eq:9}
   \end{equation}
   where $h\equiv H_0/[100\text{km/s/Mpc}]$.

@@

   If $\kappa=0$, the global density evolution can be written as
   \begin{equation}
      \rho(a)=\mathcal{D}a^{-3}+C_1a^{n},
   \end{equation}
   where $\mathcal{D}=C_2-C_1$. The conditions in which the current 
   model acts as a quintessence cosmology are given by $\mathcal{D}>0,
   C_1>0,$ and $w_{\text{IDM}}=-1-n/3$. This solution is mathematically
   equivalent to that of the gravitational matter creation model of \cite{Lima_2008}.
   The Hubble flow is now given by
   \begin{equation}
      \left(\frac{H}{H_0}\right)^2=\Omega_{2,0}a^{-3}+\Omega_{1,0}a^{n},
   \end{equation}
   where $\Omega_{2,0}=8\pi G\mathcal{D}/3H_0^2$ and 
   $\Omega_{1,0}=8\pi GC_1/3H_0^2$, respectively.(\cite{Basilakos_2009})

\section{Dataset}

@@

   It is widely known that the Hubble parameter $H(z)$ depends on
   the differential age as a function of redshift $z$ in the form
   \begin{equation}
      H(z)=-\frac{1}{1+z}\frac{\mathrm{d}z}{\mathrm{d}t},\label{eq:12}
   \end{equation}
   which provides a direct measurement on $H(z)$ based on
   $\mathrm{d}z/\mathrm{d}t$.
   OHD measurements have recently been acquired mainly employing
   cosmic chronometers (CC). The CC method is used to provide 33 observational
   data points, which are taken in the redshift range [0.07, 1.965].
   The Table \ref{tab:1} lists the OHD dataset used in this analysis.
   In this case, $\chi^2$ can be defined as
   \begin{equation}
      \chi_{\text{OHD}}^2=\sum_i\frac{(H_{\text{th}}-H_{\text{data}})^2}{\sigma_i^2}.
   \end{equation}

   \begin{table}
      \centering
      \begin{tabular}{ccl}
         \hline\hline
         $z$ & $H(z)$ & Reference \\
         \hline     
         0.07 & $69 \pm 19.6$ & \cite{Zhang_2014} \\
         0.09 & $69 \pm 12$ & \cite{PhysRevD.71.123001} \\
         0.12 & $68.6 \pm 26.2$ & \cite{Zhang_2014} \\
         0.17 & $83 \pm 8$ & \cite{PhysRevD.71.123001} \\
         0.179 & $75 \pm 4$ & \cite{M.Moresco_2012} \\
         0.199 & $75 \pm 5$ & \cite{M.Moresco_2012} \\
         0.2 & $72.9 \pm 29.6$ & \cite{Zhang_2014} \\
         0.27 & $77 \pm 14$ & \cite{PhysRevD.71.123001} \\
         0.28 & $88.8 \pm 36.6$ & \cite{Zhang_2014} \\
         0.352 & $83 \pm 14$ & \cite{M.Moresco_2012} \\
         0.3802 & $83 \pm 13.5$ & \cite{Moresco_2016}  \\
         0.4 & $95 \pm 17$ & \cite{PhysRevD.71.123001} \\
         0.4004 & $77 \pm 10.2$ & \cite{Moresco_2016} \\
         0.4247 & $87.1 \pm 11.2$ & \cite{Moresco_2016} \\
         0.4497 & $92.8 \pm 12.9$ & \cite{Moresco_2016} \\
         0.47 & $89 \pm 34$ & \cite{10.1093/mnras/stx301} \\
         0.4783 & $80.9 \pm 9$ & \cite{Moresco_2016} \\
         0.48 & $97 \pm 62$ & \cite{Daniel.Stern_2010} \\
         0.593 & $104 \pm 13$ & \cite{M.Moresco_2012} \\
         0.68 & $92 \pm 8$ & \cite{M.Moresco_2012} \\
         0.75 & $98.8 \pm 33.6$ & \cite{Borghi_2022} \\
         0.781 & $105 \pm 12$ & \cite{M.Moresco_2012} \\
         0.8 & $113.1 \pm 15.1$ & \cite{Jiao_2023} \\
         0.875 & $125 \pm 17$ & \cite{M.Moresco_2012} \\
         0.88 & $90 \pm 40$ & \cite{Daniel.Stern_2010} \\
         0.9 & $117 \pm 23$ & \cite{PhysRevD.71.123001} \\
         1.037 & $154 \pm 20$ & \cite{M.Moresco_2012} \\
         1.3 & $168 \pm 17$ & \cite{PhysRevD.71.123001} \\
         1.363 & $160 \pm 33.6$ & \cite{10.1093/mnrasl/slv037} \\
         1.43 & $177 \pm 18$ & \cite{PhysRevD.71.123001} \\
         1.53 & $140 \pm 14$ & \cite{PhysRevD.71.123001} \\
         1.75 & $202 \pm 40$ & \cite{PhysRevD.71.123001} \\
         1.965 & $186.5 \pm 50.4$ & \cite{10.1093/mnrasl/slv037} \\
         \hline    
      \end{tabular}
      \caption{The OHD dataset from different reference.}
      \label{tab:1}
   \end{table}

@@

   We use the fiducial SNe Ia magnitude ($M_b$) determined from SH0ES 2021 Cepheid 
   host distances (\cite{Riess_2022}), which gives the $\mu_{\text{data}}$ and we 
   give the $\chi^2$ as 
   \begin{equation}
      \chi_{\text{SNe}}^2=\Delta C^{-1} \Delta^{\text{T}},
   \end{equation}
   where $\Delta=(\mu_{\text{th}}-\mu_{\text{data}})$ and $C^{-1}$ is the inverse of the
   covariance matrix of the SNe Ia data, the distance modulus is $\mu=5\log_{10}(d_L/\text{Mpc})+25$, and the
   luminosity distance $d_L$ can be given as a function of redshift $z$
   \begin{equation}
      d_L=(1+z)\int_0^z\frac{c\mathrm{d}z'}{H(z')}.\label{eq:15}
   \end{equation}
   To eliminate the advanced restriction to $H_0$ from $M_b$, we adopt
   the likelihood function as
   \begin{equation}
      \widetilde{\chi}_{\text{SNe}}^2=\chi_{\text{SNe}}^2-\frac{B^2}{\mathcal{C}}+\ln\left(\frac{\mathcal{C}}{2\pi}\right),
   \end{equation}
   where $B$ is the sum of $\Delta C^{-1}$ and $\mathcal{C}$ is the sum of $C^{-1}$.

@@

   The Baryon acoustic oscillation (BAO) method provides a key cosmological probe
   sensitive to the cosmic expansion history with well-controlled systematics.
   We use two BAO datasets from the SDSS-IV (\cite{PhysRevD.103.083533}) and 
   DESI DR2 (\cite{desicollaboration2025desidr2resultsii}),
   which are given at Table \ref{tab:2} and Table \ref{tab:3}, respectively.
   The redshift is up to 2.33 both in the SDSS and the DESI 2024 dataset.

   The $\chi^2$ function for the BAO data is defined as
   \begin{equation}
      \chi_{\text{BAO}}^2=\sum_i\frac{(D_{\text{th}}/r_{\text{d}}-D_{\text{data}}/r_{\text{d}})^2}{\sigma_i^2},
   \end{equation}
   where $D$ refers to $D_{\text{M}}$, $D_{\text{H}}$, or $D_{\text{V}}$, which ara given as
   \begin{eqnarray}
      D_{\text{M}}(z)&=&c\int_0^z\frac{\text{d}z'}{H(z')},\\
      D_{\text{H}}(z)&=&\frac{c}{H(z)},\\
      D_{\text{V}}(z)&=&\left[zD_{\text{M}}^2(z)D_{\text{H}}(z)\right]^{1/3},
   \end{eqnarray}
   and $r_{\text{d}}$ is the sound horizon at the drag epoch, which is given as
   \begin{equation}
      r_{\text{d}}=\int_{z_{\text{drag}}}^{\infty}\frac{c_s(z')\text{d}z'}{H(z')},
   \end{equation}
   where $c_s$ is the sound speed prior to recombination.

@@

   \begin{table*}
      \centering
      \begin{tabular}{lcccc}
         \hline\hline
         Tracer & $z_{\text{eff}}$ & $D_{\text{V}}/r_{\text{d}}$ & 
         $D_{\text{M}}/r_{\text{d}}$ & $D_{\text{H}}/r_{\text{d}}$ \\
         \hline
         MGS & 0.15 & $4.51 \pm 0.14$ & --- & --- \\
         BOSS galaxy & 0.38 & --- & $10.27 \pm 0.15$ & $24.89 \pm 0.58$ \\
         BOSS galaxy & 0.51 & --- & $13.38 \pm 0.18$ & $22.43 \pm 0.48$ \\
         eBOSS LRG & 0.70 & --- & $17.65 \pm 0.30$ & $19.78 \pm 0.46$ \\
         eBOSS ELG & 0.85 & --- & $19.5 \pm 1.0$ & $19.6 \pm 2.1$ \\
         eBOSS quasar& 1.48 & --- & $30.21 \pm 0.79$ & $13.23 \pm 0.47$ \\
         Ly$\alpha$-Ly$\alpha$ & 2.33 & --- & $37.6 \pm 1.9$ & $8.93 \pm 0.28$ \\
         Ly$\alpha$-quasar & 2.33 & --- & $37.3 \pm 1.7$ & $9.08 \pm 0.34$ \\
         \hline
      \end{tabular}
      \caption{The BAO+RSD measurements dataset from SDSS-IV.}
      \label{tab:2}
   \end{table*}

   \begin{table*}
      \centering
      \begin{tabular}{lcccc}
         \hline\hline
         Tracer & $z_{\text{eff}}$ & $D_{\text{V}}/r_{\text{d}}$ & 
         $D_{\text{M}}/r_{\text{d}}$ & $D_{\text{H}}/r_{\text{d}}$ \\
         \hline
         BGS & 0.295 & $7.942 \pm 0.075$  & --- & --- \\
         LRG1 & 0.510 & $12.720 \pm 0.099$ & $13.588 \pm 0.167$ & $21.863 \pm 0.425$ \\
         LRG2 & 0.706 & $16.050 \pm 0.110$ & $17.351 \pm 0.177$ & $19.455 \pm 0.330$ \\
         LRG3+ELG1 & 0.934 & $19.721 \pm 0.091$ & $21.576 \pm 0.152$ & $17.641 \pm 0.193$ \\
         ELG2 & 1.321 & $24.252 \pm 0.174$ & $27.601 \pm 0.318$ & $14.176 \pm 0.221$ \\
         QSO & 1.484 & $26.055 \pm 0.398$ & $30.512 \pm 0.760$ & $12.817 \pm 0.516$ \\
         Ly$\alpha$ & 2.330 & $31.267 \pm 0.256$ & $38.988 \pm 0.531$ & $8.632 \pm 0.101$ \\
         \hline
         LRG3 & 0.922 & $19.656 \pm 0.105$ & $21.648 \pm 0.178$ & $17.577 \pm 0.213$ \\
         ELG1 & 0.955 & $20.008 \pm 0.183$ & $21.707 \pm 0.335$ & $17.803 \pm 0.297$ \\    
         \hline
      \end{tabular}
      \caption{The BAO dataset from DESI DR2.}
      \label{tab:3}
   \end{table*}

\section{Constraint results}

   We use the Markov chain Monte Carlo (MCMC) method based 
   on the opening package \texttt{emcee} to give a global constraints
   to the free parameters $\Omega_{2,0}$ and $\log_{10}(\kappa C_1)$ in 
   Model 1 and $n$ in Model 2.
   Besides, we also add the parameter $H_0$ or $h$ to give the constraints
   to the dark matter particle mass $M_x$.
   The prior range set for free parameters are given at Table \ref{tab:4}

   \begin{table}
      \centering
      \begin{tabular}{lll}
         \hline\hline
         parameter & initial & prior \\
         \hline
         $\Omega_{2,0}$ & 0.3 & $\mathcal{U}[0.0,1.0]$ \\
         $n$ & 0 & $\mathcal{U}[-10,10]$ \\
         $\log_{10}(\kappa C_1/$Gyr${}^{-1})$ & --- & $\mathcal{U}[-10,0]$ \\
         $H_0$[km/s/Mpc] & 70 & $\mathcal{U}[60,80]$ \\
         \hline
         $r_{\text{d}}h$[Mpc] & 100 & $\mathcal{U}[50,150]$ \\
         \hline
      \end{tabular}
      \caption{Parameters and priors used in analysis. 
      All of the priors are flat in the ranges given. 
      The $\log_{10}(\kappa C_1)$ does not have a predicted value, 
      and the initial value of each chain is selected randomly uniformly
      within the prior range.}
      \label{tab:4}
   \end{table}

   For the dataset which cannot give a constraint to $H_0$, 
   we use $h \simeq 0.7$ as the default value to calculate the 
   dark matter particle mass $M_x$

@@

   The Eq.(\ref{eq:8}) convergant to the flat ΛCDM model as
   $\log_{10}(\kappa C_1)\to-\infty$, thus the constraints
   could only give the upper limit of what, and give the lower limit
   of $M_x$ according to Eq.(\ref{eq:9}), respectively.

>>>>

\section{Introduction}\label{sec:1}

@@

   In this paper, we use the newly revised data from $H(z)$, supernovae Ia 
   and baryon acoustic oscillation by using the Markov chain Monte
   Carlo (MCMC) method. The constraint result gives an upper limit for the 
   linear ratio between the dark matter annihilation cross section $\langle\sigma v\rangle$
   and the dark matter particle mass $M_{\chi}$ about $9\times10^{-15} \;\text{cm}^{3}\text{s}^{-1}\text{GeV}^{-1}$.

   The paper is arranged as follows: In Sec.\ref{sec:2}, we review the basic equations in the IDM model,
   in Sec.\ref{sec:3}, we introduce the observational datasets used in this analysis, in Sec.\ref{sec:4}, 
   we give the constraint results of the IDM model and compare our results with 
   other relevant dark matter constraints, and finally we conclude in Sec.\ref{sec:5}.

\section{The basic equations in the IDM model}\label{sec:2}

   We assume that the total density of the cosmic fluid obeys
   the collisional Boltzmann equation (\cite{Kolb:1990vq})
   \begin{equation}
      \dot{\rho}+3H\rho+\kappa\rho^2-2\Psi=0,
      \label{eq:1}
   \end{equation}
   where $\rho$ is the total energy-density of the cosmic fluid,
   containing dark matter, baryons, and any type of exotic energy,
   $\Psi$ is the rate of creation of DM particle pairs, and the
   annihilation parameter $\kappa(\geq0)$ is given by:
   \begin{equation}
      \kappa=\frac{\langle\sigma v\rangle}{M_{\chi}},
      \label{eq:2}
   \end{equation}
   where $\sigma$ is the cross section for annihilation, $v$ is
   the mean particle velocity, and $M_{\chi}$ is the mass of the DM
   particle.
   Compared to the usual fluid equation, the effective pressure term
   is \begin{equation}
      P=\frac{\kappa\rho^2-\Psi}{3H}.
      \label{eq:3}
   \end{equation}
   When $\kappa\rho^2-\Psi<0,$ what means that the IDM particle
   creation term is larger than the annihilation item, IDM may serve
   as a negative pressure source in the global dynamics of the Universe,
   like the role of Dark Energy in the general cosmological models.

   \cite{Basilakos_2009} identified two functional forms for which
   the previous Boltzmann equation can be solved analytically.
   Refering to Appendix B in \cite{Basilakos_2009}, only one of these two is of interest
   because it provides a ``$\propto\alpha^{-3}$'' dependence of the scale factor,
   which is \begin{equation}
      \Psi(\alpha)=\alpha H(\alpha)R(\alpha)=C_1(n+3)\alpha^nH(\alpha)+\kappa C_1^2 \alpha^{2m}.
      \label{eq:4}
   \end{equation}
   And the total energy density is
   \begin{equation}
      \rho(\alpha)=C_1 \alpha^n+\frac{\alpha^{-3}F(\alpha)}{C_2-\int_1^{\alpha}x^{-3}f(x)F(x)\mathrm{d}x},
      \label{eq:5}
   \end{equation}
   where $f(\alpha)=-\kappa/[\alpha H(\alpha)],$ and the kernal function $F(\alpha)$ has
   the form \begin{equation}
      F(\alpha)=\exp\left[-2\kappa C_1\int_1^{\alpha}\frac{x^{n-1}}{H(x)}\mathrm{d}x\right].
      \label{eq:6}
   \end{equation}
   The first term of Eq.(\ref{eq:5}) is the density corresponding to the
   residual matter creation that results from a possible disequilibrium
   between the particle creation and annihilation processes, while the second
   term can be viewed as the energy density of the self-IDM particles that are
   dominated by the annihilation process.

@@

   If $n=0$, the global density evolution can be transformed as
   \begin{equation}
      \rho(\alpha)=C_1+\alpha^{-3}\frac{e^{-2\kappa C_1(t-t_0)}}{C_2+\kappa Z(t)},
      \label{eq:7}
   \end{equation}
   where $Z(t)=\int_{t_0}^t \alpha^{-3}e^{-2\kappa C_1(t'-t_0)}\mathrm{d}t'$ (\cite{Basilakos_2009}).
   Using the usual unit-less $\Omega$-like parameterization, we obtain that
   \begin{equation}
      \left(\frac{H}{H_0}\right)^2=\Omega_{1,0}+\frac{\Omega_{1,0}\Omega_{2,0}\alpha^{-3}e^{-2\kappa C_1(t-t_0)}}
      {\Omega_{1,0}+\kappa C_1\Omega_{2,0}Z(t)},
      \label{eq:8}
   \end{equation}
   where $\Omega_{1,0}=8\pi GC_1/3H_0^2$ and $\Omega_{2,0}=8\pi G/3H_0^2C_2$,
   which related to $\Omega_{\Lambda}$ and $\Omega_m$ in the $\Lambda$CDM model,
   respectively.
   From Eq.({\ref{eq:2}}), we can also give the mass of the DM particle
   related to the range of $\kappa C_1$(in the unit of Gyr${}^{-1}$)
   \begin{equation}
      M_{\chi}=\frac{3.325\times10^{-12}}{\kappa C_1}
      \frac{\langle\sigma v\rangle}{10^{-23}}h^2(1-\Omega_{2,0})\,\text{GeV},
      \label{eq:9}
   \end{equation}
   where $h\equiv H_0/[100\text{km/s/Mpc}]$.

@@

   If $n\neq0$, the global density evolution is Eq.(\ref{eq:5}), and the 
   usual unit-less $\Omega$-like can be written as
   \begin{equation}
      \left(\frac{H}{H_0}\right)^2 = \Omega_{1,0} \alpha^n + 
      \frac{\Omega_{1,0} \Omega_{2,0} \alpha^{-3}F(\alpha)}
      {\Omega_{1,0} + \kappa C_1 \Omega_{2,0} \int_1^\alpha 
      \frac{F(x)}{x^4 H(x)} \mathrm{d}x},
      \label{eq:10}
   \end{equation}
   where $\Omega_{1,0}$ and $\Omega_{2,0}$ are the same defined as Eq.(\ref{eq:8}), and 
   $F(\alpha)$ is given by Eq.(\ref{eq:6}).

   Considering the limit of $\kappa \to0$, the global density evolution can be
   simplified as
   \begin{equation}
      \rho(\alpha)=C_1 a^{n}+\frac{1}{C_2} a^{-3},
      \label{eq:11}
   \end{equation}
   and Eq.(\ref{eq:10}) becomes
   \begin{equation}
      \left(\frac{H}{H_0}\right)^2=\Omega_{1,0}\alpha^{n}+\Omega_{2,0}\alpha^{-3}.
      \label{eq:12}
   \end{equation}
   The conditions $n>2$ indicates that the IDM model has an inflection point (\cite{Basilakos_2009}),
   and we define $w_{\text{IDM}}=-1-n/3$ as the equation of state (EoS) of the IDM model.
   This solution is mathematically equivalent to that of the gravitational matter 
   creation model of \cite{Lima_2008}.

\section{Dataset}\label{sec:3}

@@

   It is widely known that the Hubble parameter $H(z)$ depends on
   the differential age as a function of redshift $z$ in the form
   \begin{equation}
      H(z)=-\frac{1}{1+z}\frac{\mathrm{d}z}{\mathrm{d}t},
      \label{eq:13}
   \end{equation}
   which provides a direct measurement on $H(z)$ based on
   $\mathrm{d}z/\mathrm{d}t$.
   OHD measurements have recently been acquired mainly employing
   cosmic chronometers (CC). The CC method is used to provide 33 observational
   data points, which are taken in the redshift range [0.07, 1.965].
   The Table \ref{tab:1} lists the OHD dataset used in this analysis.
   In this case, $\chi^2$ can be defined as
   \begin{equation}
      \chi_{\text{OHD}}^2=\sum_i\frac{(H_{\text{th}}-H_{\text{data}})^2}{\sigma_i^2}.
      \label{eq:14}
   \end{equation}

   \begin{table}
      \centering
      \begin{tabular}{ccl}
         \hline\hline
         $z$ & $H(z)$ & Reference \\
         \hline     
         0.07 & $69 \pm 19.6$ & \cite{Zhang_2014} \\
         0.09 & $69 \pm 12$ & \cite{PhysRevD.71.123001} \\
         0.12 & $68.6 \pm 26.2$ & \cite{Zhang_2014} \\
         0.17 & $83 \pm 8$ & \cite{PhysRevD.71.123001} \\
         0.179 & $75 \pm 4$ & \cite{M.Moresco_2012} \\
         0.199 & $75 \pm 5$ & \cite{M.Moresco_2012} \\
         0.2 & $72.9 \pm 29.6$ & \cite{Zhang_2014} \\
         0.27 & $77 \pm 14$ & \cite{PhysRevD.71.123001} \\
         0.28 & $88.8 \pm 36.6$ & \cite{Zhang_2014} \\
         0.352 & $83 \pm 14$ & \cite{M.Moresco_2012} \\
         0.3802 & $83 \pm 13.5$ & \cite{Moresco_2016}  \\
         0.4 & $95 \pm 17$ & \cite{PhysRevD.71.123001} \\
         0.4004 & $77 \pm 10.2$ & \cite{Moresco_2016} \\
         0.4247 & $87.1 \pm 11.2$ & \cite{Moresco_2016} \\
         0.4497 & $92.8 \pm 12.9$ & \cite{Moresco_2016} \\
         0.47 & $89 \pm 34$ & \cite{10.1093/mnras/stx301} \\
         0.4783 & $80.9 \pm 9$ & \cite{Moresco_2016} \\
         0.48 & $97 \pm 62$ & \cite{Daniel.Stern_2010} \\
         0.593 & $104 \pm 13$ & \cite{M.Moresco_2012} \\
         0.68 & $92 \pm 8$ & \cite{M.Moresco_2012} \\
         0.75 & $98.8 \pm 33.6$ & \cite{Borghi_2022} \\
         0.781 & $105 \pm 12$ & \cite{M.Moresco_2012} \\
         0.8 & $113.1 \pm 15.1$ & \cite{Jiao_2023} \\
         0.875 & $125 \pm 17$ & \cite{M.Moresco_2012} \\
         0.88 & $90 \pm 40$ & \cite{Daniel.Stern_2010} \\
         0.9 & $117 \pm 23$ & \cite{PhysRevD.71.123001} \\
         1.037 & $154 \pm 20$ & \cite{M.Moresco_2012} \\
         1.3 & $168 \pm 17$ & \cite{PhysRevD.71.123001} \\
         1.363 & $160 \pm 33.6$ & \cite{10.1093/mnrasl/slv037} \\
         1.43 & $177 \pm 18$ & \cite{PhysRevD.71.123001} \\
         1.53 & $140 \pm 14$ & \cite{PhysRevD.71.123001} \\
         1.75 & $202 \pm 40$ & \cite{PhysRevD.71.123001} \\
         1.965 & $186.5 \pm 50.4$ & \cite{10.1093/mnrasl/slv037} \\
         \hline\hline    
      \end{tabular}
      \caption{The OHD dataset from different reference.}
      \label{tab:1}
   \end{table}

@@

   We use the fiducial SNe Ia magnitude ($M_b$) determined from SH0ES 2021 Cepheid 
   host distances (\cite{Riess_2022}), which gives the $\mu_{\text{data}}$ and we 
   give the $\chi^2$ as 
   \begin{equation}
      \chi_{\text{SNe}}^2=\Delta C^{-1} \Delta^{\text{T}},
      \label{eq:15}
   \end{equation}
   where $\Delta=(\mu_{\text{th}}-\mu_{\text{data}})$ and $C^{-1}$ is the inverse of the
   covariance matrix of the SNe Ia data, the distance modulus is $\mu=5\log_{10}(d_L/\text{Mpc})+25$, and the
   luminosity distance $d_L$ can be given as a function of redshift $z$
   \begin{equation}
      d_L=(1+z)\int_0^z\frac{c\mathrm{d}z'}{H(z')}.
      \label{eq:16}
   \end{equation}
   To eliminate the advanced restriction to $H_0$ from $M_b$, we adopt
   the likelihood function as
   \begin{equation}
      \widetilde{\chi}_{\text{SNe}}^2=\chi_{\text{SNe}}^2-\frac{B^2}{\mathcal{C}}+\ln\left(\frac{\mathcal{C}}{2\pi}\right),
      \label{eq:17}
   \end{equation}
   where $B$ is the sum of $\Delta C^{-1}$ and $\mathcal{C}$ is the sum of $C^{-1}$.

@@

   The Baryon acoustic oscillation (BAO) method provides a key cosmological probe
   sensitive to the cosmic expansion history with well-controlled systematics.
   We use two BAO datasets from the SDSS-IV (\cite{PhysRevD.103.083533}) and 
   DESI DR2 (\cite{desicollaboration2025desidr2resultsii}),
   which are given at Table \ref{tab:2} respectively.
   The redshift is up to 2.33 both in the SDSS and the DESI 2024 dataset.

   The $\chi^2$ function for the BAO data is defined as
   \begin{equation}
      \chi_{\text{BAO}}^2=\sum_i\frac{(D_{\text{th}}/r_{\text{d}}-D_{\text{data}}/r_{\text{d}})^2}{\sigma_i^2},
      \label{eq:18}
   \end{equation}
   where $D$ refers to $D_{\text{M}}$, $D_{\text{H}}$, or $D_{\text{V}}$, which ara given as
   \begin{eqnarray}
      D_{\text{M}}(z)&=&c\int_0^z\frac{\text{d}z'}{H(z')},\\
      D_{\text{H}}(z)&=&\frac{c}{H(z)},\\
      D_{\text{V}}(z)&=&\left[zD_{\text{M}}^2(z)D_{\text{H}}(z)\right]^{1/3},
      \label{eq:19-21}
   \end{eqnarray}
   and $r_{\text{d}}$ is the sound horizon at the drag epoch, which is given as
   \begin{equation}
      r_{\text{d}}=\int_{z_{\text{drag}}}^{\infty}\frac{c_s(z')\text{d}z'}{H(z')},
      \label{eq:22}
   \end{equation}
   where $c_s$ is the sound speed prior to recombination.

@@

   \begin{table*}
      \centering
      \begin{tabular}{l|l|cccc}
         \hline\hline
         Dataset & Tracer & $z_{\text{eff}}$ & $D_{\text{V}}/r_{\text{d}}$ & 
         $D_{\text{M}}/r_{\text{d}}$ & $D_{\text{H}}/r_{\text{d}}$ \\
         \hline
         \textbf{SDSS-IV} & MGS & 0.15 & $4.51 \pm 0.14$ & --- & --- \\
                      & BOSS galaxy & 0.38 & --- & $10.27 \pm 0.15$ & $24.89 \pm 0.58$ \\
                      & BOSS galaxy & 0.51 & --- & $13.38 \pm 0.18$ & $22.43 \pm 0.48$ \\
                      & eBOSS LRG & 0.70 & --- & $17.65 \pm 0.30$ & $19.78 \pm 0.46$ \\
                      & eBOSS ELG & 0.85 & --- & $19.5 \pm 1.0$ & $19.6 \pm 2.1$ \\
                      & eBOSS quasar& 1.48 & --- & $30.21 \pm 0.79$ & $13.23 \pm 0.47$ \\
                      & Ly$\alpha$-Ly$\alpha$ & 2.33 & --- & $37.6 \pm 1.9$ & $8.93 \pm 0.28$ \\
                      & Ly$\alpha$-quasar & 2.33 & --- & $37.3 \pm 1.7$ & $9.08 \pm 0.34$ \\
         \hline\hline
         \textbf{DESI DR2} & BGS & 0.295 & $7.942 \pm 0.075$  & --- & --- \\
                           & LRG1 & 0.510 & $12.720 \pm 0.099$ & $13.588 \pm 0.167$ & $21.863 \pm 0.425$ \\
                           & LRG2 & 0.706 & $16.050 \pm 0.110$ & $17.351 \pm 0.177$ & $19.455 \pm 0.330$ \\
                           & LRG3+ELG1 & 0.934 & $19.721 \pm 0.091$ & $21.576 \pm 0.152$ & $17.641 \pm 0.193$ \\
                           & ELG2 & 1.321 & $24.252 \pm 0.174$ & $27.601 \pm 0.318$ & $14.176 \pm 0.221$ \\
                           & QSO & 1.484 & $26.055 \pm 0.398$ & $30.512 \pm 0.760$ & $12.817 \pm 0.516$ \\
                           & Ly$\alpha$ & 2.330 & $31.267 \pm 0.256$ & $38.988 \pm 0.531$ & $8.632 \pm 0.101$ \\
         \hline
                           & LRG3 & 0.922 & $19.656 \pm 0.105$ & $21.648 \pm 0.178$ & $17.577 \pm 0.213$ \\
                           & ELG1 & 0.955 & $20.008 \pm 0.183$ & $21.707 \pm 0.335$ & $17.803 \pm 0.297$ \\    
         \hline\hline
      \end{tabular}
      \caption{The BAO+RSD measurements dataset from SDSS-IV and DESI DR2.}
      \label{tab:2}
   \end{table*}

\section{Constraint results}\label{sec:4}

   We use the Markov chain Monte Carlo (MCMC) method based 
   on the opening package \texttt{emcee} to give a global constraints
   to the free parameters $\Omega_{2,0}$ and $\log_{10}(\kappa C_1)$ in 
   Model 1 and $n$ in Model 2.
   Besides, we also add the parameter $H_0$ or $h$ to give the constraints
   to the dark matter particle mass $M_{\chi}$.
   The prior range set for free parameters are given at Table \ref{tab:3}

   \begin{table}
      \centering
      \begin{tabular}{lll}
         \hline\hline
         parameter & initial & prior \\
         \hline
         $\Omega_{2,0}$ & 0.3 & $\mathcal{U}[0.0,1.0]$ \\
         $n$ & 0 & $\mathcal{U}[-5,5]$ \\
         $\log_{10}(\kappa C_1/$Gyr${}^{-1})$ & --- & $\mathcal{U}[-10,0]$ \\
         $H_0$[km/s/Mpc] & 70 & $\mathcal{U}[60,80]$ \\
         \hline
         $r_{\text{d}}h$[Mpc] & 100 & $\mathcal{U}[50,150]$ \\
         \hline
      \end{tabular}
      \caption{Parameters and priors used in analysis. 
      All of the priors are flat in the ranges given. 
      The $\log_{10}(\kappa C_1)$ does not have a predicted value, 
      and the initial value of each chain is selected randomly uniformly
      within the prior range.}
      \label{tab:3}
   \end{table}

   For the dataset which cannot give a constraint to $H_0$, 
   we use $h \simeq 0.7$ as the default value to calculate the 
   annihilation parameter $\kappa$ and dark matter particle mass $M_{\chi}$.

@@

   The Eq.(\ref{eq:8}) convergant to the flat ΛCDM model as
   $\log_{10}(\kappa C_1)\to-\infty$, thus the constraints
   could only give the upper limit of what, and give the lower limit
   of $M_{\chi}$ according to Eq.(\ref{eq:9}), respectively. We 
   assume the dark matter annihilation cross section in IDM model
   a constant that $\langle\sigma v\rangle \simeq 10^{-23}\text{cm}^{3}\text{s}^{-1}$.
